---
title: "CINSignatureQuantification"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{CINSignatureQuantification}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(CINSignatureQuantification)
```

An example data set is provided in this package. This data is a `data.frame` containing segmented copy number data for 478 samples with detectable chromosomal instability (CIN).

```{r load data}
data("TCGA_478_Samples_SNP6_GOLD")
head(TCGA_478_Samples_SNP6_GOLD)
```
## Load data

The first step is to load the data into the `CNQuant` object. This is a S4 class object which standardises the input and additional meta data for downstream analysis. Here the `data` argument can be a loaded R data.frame object (as shown in this vignette), a file path to a delimited file, or a `QDNAseqCopyNumbers` object from [QDNASeq](https://github.com/ccagc/QDNAseq) which has been segmented. A name for the experiment and the genome build to use are also specified. Viewing this object will print details of the contained data.

```{r createCNQuant}
cnobj <- createCNQuant(data = TCGA_478_Samples_SNP6_GOLD,experimentName = "VignetteExample",build = "hg19")
cnobj
```
### CNQuant object

Additional meta data can be retrieved with `getExperiment()`

```{r getExp}
getExperiment(cnobj)
```

The `CNQuant` object contains slots for all the required prerequisite data for copy number signature analysis. Of these, 

```{r CNQuant_segs}
# Returns the original segment table
segmentTable <- getSegments(cnobj)
head(segmentTable)
```

```{r CNQuant_samples}
# Returns all sample ids 
getSamples(cnobj)[1:10]
```

```{r CNQuant_sampleFeatures}
# Returns sample features computed on CNQuant initialisation
Samplefeatures <- getSamplefeatures(cnobj)
head(Samplefeatures)
```

If you have additional sample features or clinical data you can add those to the object using the `addsampleFeatures()` function.

```{r CNQuant_sampleFeatures2}
data(test.sample.features)
head(test.sample.features)
```

```{r CNQuant_sampleFeatures3}
cnobj <- addsampleFeatures(object = cnobj,sample.data = test.sample.features,id.col = "sample")
NewSamplefeatures <- getSamplefeatures(cnobj)
head(NewSamplefeatures)
```

Sub setting of `CNQuant` objects is implemented using the native R bracket `[` notation. This can be performed using either numerical indexing or specifying available sample identifiers. This sub setting functions at any point in the analysis pipeline.

```{r subsetting1}
cnobj[1:10]
```

```{r subsetting2}
cnobj[c(1,24,5,77,100)]
```

```{r subsetting2}
cnobj[getSamples(cnobj)[1:10]]
```

For the sake of computational efficiency, we will subset the full data set to 20 samples.

```{r vignette_dataset}
cnobj <- cnobj[1:20]
cnobj
```

## Calculate features

Feature distributions can be calculate using the `calculateFeatures()` function. 

```{r calcFeats}
cnobj <- calculateFeatures(object = cnobj,method = "drews",cores = 1)
cnobj
```

These features can be retrieved as follows

```{r getFeats}
# Returns list of data.frames containing the computed copy number features
feats <- getFeatures(cnobj)
head(feats[[1]])
```
## Calculate Sample-by-component

```{r calcSxC}
cnobj <- calculateSampleByComponentMatrix(object = cnobj)
cnobj
```

The sample-by-component matrix can be retrieved as follows

```{r getSxC}
# Returns the computed sample-by-component matrix
SxC <- getSampleByComponent(cnobj)
head(SxC)
```

## Calculate signature activites

```{r calcSxC}
cnobj <- calculateActivity(object = cnobj)
cnobj
```

The sample-by-signature activity matrix can be retrieved as follows

```{r getSxC}
# Returns the computed sample-by-component matrix
sigAct <- getActivities(cnobj,type = "threshold")
head(sigAct)
```

## One-click function

A wrapper function is provided to allow the entire pipeline to run as a singe command. Here the output object is a `SigQuant` object with all slots and calculations performed.

```{r quantSub}
# Subset samples from the total dataset
subsample <- as.character(sample(TCGA_478_Samples_SNP6_GOLD$sample,size = 20))
TCGA_478_Samples_SNP6_GOLD_subset <- TCGA_478_Samples_SNP6_GOLD[TCGA_478_Samples_SNP6_GOLD$sample %in% subsample,]
```

```{r quantFull}
cnobj <- quantifyCNSignatures(object = TCGA_478_Samples_SNP6_GOLD_subset,
                              experimentName = "VignetteExample",
                              method = "drews",
                              cores = 1,
                              build = "hg19")
cnobj
```

## Clinical predictors

Clinical predictor functions are provided to allow for binary classification of samples based on signature activities. This includes the platinum-based treatment response predictor included in the associated publication and a function to construct custom classifications.

```{r clinPt}
clinPredictionPlatinum(object = cnobj)
```

```{r clinDenov}
# Not a real predictor
clinPredictionDenovo(cnobj,
             sampTrain = getSamples(cnobj[1:5]),
             sigsTrain = c("CX1","CX6"))
```
## Plotting functions

```{r segmentPlots}
# Sample can be specified as a numerical index or sample name
plotSegments(object = cnobj,sample = 1,cn.max = 8)
```

```{r SxCPlots}
# Additional arguments can be provided for the heatmap() function, such as "main" shown in the example below
plotSampleByComponent(object = cnobj,main="VignetteExample")
```

```{r activitiesPlots}
# custom colours for signatures can be provided with the cols argument, the length of which must match the number of signatures
plotActivities(object = cnobj,type = "threshold")
```

```{r sessionInfo}
sessionInfo()
```
