---
title: "CINSignaturesQuantification_vignette"

---
title: "CINSignatureQuantification - Simple and quick measuring of copy number signatures in cancers"
author: "Philip Smith <cre>, Ruben Drews <cre, auth>, University of Cambridge <cph, fnd>"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{CINSignaturesQuantification_vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
#knitr::opts_chunk$set(echo = TRUE)
```

```{r load_lib}
#library(CINSignatureQuantification)
```

```{r test_data}
#test_data = readRDS("../inst//TCGA_478_Samples_SNP6_GOLD.rds")
#head(test_data)
```

### Load data

```{r load_data}
#myData = createCignatures(data = test_data,experimentName = "TCGA/PCAWG overlap with dCIN")
#myData
```

### Class features

#### experiment class slot

```{r exp_data}
#getExperiment(object = myData)
```

#### Samples

```{r get_samples}
#samples = getSamples(object = myData)
#samples[1:10]
```

#### Segment data

```{r get_segs}
#segs = getSegments(object = myData)
#head(segs)
```
#### Sample features

```{r get_sampleFeats}
#sampleFeats = getSamplefeatures(object = myData)
#head(sampleFeats)
```

```{r add_sampleFeats}
#more_sample_data = read.table("data/test_sample_feature_data.tsv",
#                               header = TRUE,
 #                              sep="\t")
#head(more_sample_data)
#myData = addsampleFeatures(object = myData,sample.data = more_sample_data,id.col = "sample")
#sampleFeats = getSamplefeatures(object = myData)
#head(sampleFeats)
```
### Object subsetting

```{r subsetting_data}
#myData[1:10]
```

```{r subsetting_data2}
#subSamples = getSamples(myData)[1:20]
#myData[subSamples]
#getExperiment(myData[subSamples])
```

### CN features

```{r calc_features}
#myData
#myData = calculateFeatures(object = myData,method = "mac")
#myData
#getExperiment(myData)
```

```{r subsetting_data3}
#myData = myData[1:100]
#myData
#length(unique(myData@featData$segsize$I))
```

### SamplexComponent matrix
```{r sampCompMat}
#myData = calculateSampleByComponentMatrix(object = myData)
#myData
```

```{r sampCompMat_details}
#names(myData@featFitting)
#myData@featFitting$method
#myData@featFitting$model
```

```{r subsetting_data4}
# = myData[1:50]
#myData
```

```{r sxc_plotting}
#lotSampleByComponent(object = myData)
```

